---
title: "12 - Compute Tree Canopy"
author: "Sofia Navarro"
date: "28/02/2021" 
output:
  html_document:
    collapsed: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Task 1: Compute tree density by neighborhood (I)
In order to compute tree density by neighborhood you need two things. You will need to know the area of the neighborhoods, which you will compute in the next exercise. And you need the **tree counts by neighborhood** which is the focus of this exercise.

You will produce counts of all trees by neighborhood in NYC and create a single data frame with a column for total trees. The result should be a data frame with **no geometry**.


```{r count-trees, eval=FALSE, echo=TRUE}
# Load data and libraries
library(sf)
library(raster)
library(dplyr)

# Read in the trees shapefile
trees <- st_read("data/trees.shp")

# Read in the neighborhood shapefile
neighborhoods <- st_read("data/neighborhoods.shp")

# Read in the tree canopy single-band raster
canopy <- raster("data/canopy.tif")

# Read in the manhattan Landsat image multi-band raster
manhattan <- brick("data/manhattan.tif")

# Compute the counts of all trees by hood
tree_counts <- count(trees, boroname)

# Take a quick look
head(tree_counts)

# Remove the geometry
tree_counts_no_geom <- st_set_geometry(tree_counts, NULL)

# Rename the n variable to tree_cnt
tree_counts_renamed <- rename(tree_counts_no_geom, tree_cnt = n)
head(tree_counts_renamed)  

head(tree_counts_renamed)

# Create histograms of the total counts
hist(tree_counts_renamed$tree_cnt)
```

Great, you've completed the first step toward computing tree densities. You now have tree counts by neighborhoods and in the next exercise you'll compute the neighborhood areas.

# Task 2: Compute tree density by neighborhood (II)
We have the tree counts (from the previous exercise). In this exercise you will compute neighborhood areas, add them to the neighborhood `sf` object and then you'll join in the non-spatial tree counts data frame from the previous exercise.


```{r tree-density, eval=FALSE, echo=T}
# Compute areas and unclass
areas <- unclass(st_area(neighborhoods))

# Add the areas to the neighborhoods object
neighborhoods_area <- mutate(neighborhoods, area = areas)

head(neighborhoods_area)

# Join neighborhoods and counts by shared columns (find the matching one in neighborhoods)
neighborhoods_counts <- left_join(neighborhoods_area, 
                            tree_counts_renamed, by = c("boro_name" = "boroname"))

# Replace NA values with 0
neighborhoods_counts <- mutate(neighborhoods_counts, 
                            tree_cnt = ifelse(is.na(tree_cnt), 
                                              0, tree_cnt))
# Compute the density
neighborhoods_counts <- mutate(neighborhoods_counts, 
                               tree_density = tree_cnt/area)

head(neighborhoods_counts)
```

You're part way there. You have now computed the tree density variable using, in part, the `sf` function `st_area()`. In the next exercises you will compute the average tree canopy values and then compare.

# Task 3: Compute average tree canopy by neighborhood
In the previous exercises you computed tree density by neighborhood using tree counts. In this exercise you will compute average tree canopy by neighborhood as a percentage so that we can compare if the results are similar.



```{r tree-canopy, eval=FALSE, echo=TRUE}
# Confirm that you have the neighborhood density results
head(neighborhoods_counts)

# Transform the neighborhoods CRS to match the canopy layer
neighborhoods_crs <- st_transform(neighborhoods_counts, crs = crs(canopy, asText = TRUE))

# Convert neighborhoods object to a Spatial object (optional)
neighborhoods_sp <- as(neighborhoods_crs, "Spatial")

# Compute the mean of canopy values by neighborhood 
canopy_neighborhoods <- raster::extract(canopy, neighborhoods_sp, fun = mean)

# Add the mean canopy values to neighborhoods
neighborhoods_avg_canopy <- mutate(neighborhoods, avg_canopy = canopy_neighborhoods)
```

Excellent! Note that you transformed the neighborhoods object's CRS. This is actually not strictly necessary because `extract()` can transform CRS on the fly. But it will be needed for plotting and other operations later so doing manually is important here.

# Task 4: Create plots using ggplot2
As an initial review of the data you created you will compute correlations and create histograms and a scatter plot using `ggplot2`.


```{r ggplot, eval=FALSE, echo=TRUE}
# Load the ggplot2 package
library(ggplot2)

# Create a histogram of tree density (tree_density)
ggplot(neighborhoods_counts, aes(x = tree_density)) + 
  geom_histogram(color = "white")

head(neighborhoods_avg_canopy)

# Create a histogram of average canopy (avg_canopy)
ggplot(neighborhoods_avg_canopy, aes(x = avg_canopy)) + 
  geom_histogram(color = "white")

# Make neighborhoods_avg_canopy into sf object
convert <- as(neighborhoods_avg_canopy, "sf")

# Join neighborhoods_counts and neighborhoods_avg_canopy together
neighborhoods_combined <- st_join(neighborhoods_counts, convert, by = c("boro_name" = "boro_name"))

# Create a scatter plot of tree_density vs avg_canopy
ggplot(neighborhoods_combined, aes(x = tree_density, y = avg_canopy)) + 
    geom_point() + 
    stat_smooth()

# Compute the correlation between density and canopy
cor(neighborhoods_combined$tree_density, neighborhoods_combined$avg_canopy)
```


Nice! Ggplot2 is great for creating data graphics and in the next exercise you'll see that you can use ggplot2 to make maps. In this case the scatter plot and correlation suggest an unexpected result. If anything, the street tree density data and tree canopy data are negatively correlated. You will confirm this with maps in the next tasks and you will learn why.

# Task 5: Create a map using ggplot2
The `geom_sf()` function operates like any other layer in `ggplot2` where you can link variables to aesthetics on the plot through the `aes()` function. In a mapping context this might mean, for example, creating a choropleth map by color coding the polygons based on a variable. If you leave off the aesthetic mapping `geom_sf()` will map the geometry alone.

`devtools::install_github("tidyverse/ggplot2")`


```{r map, eval=FALSE, echo=TRUE}
# Simplify name
neighborhoods <- neighborhoods_combined

# Plot the tree density with default colors
ggplot(neighborhoods) + 
  geom_sf(aes(fill = tree_density))

# Plot the tree canopy with default colors
ggplot(neighborhoods) + 
  geom_sf(aes(fill = avg_canopy))
  
# Plot the tree density using scale_fill_gradient()
ggplot(neighborhoods) + 
  geom_sf(aes(fill = tree_density)) + 
  scale_fill_gradient(low = "#edf8e9", high = "#005a32")

# Plot the tree canopy using the scale_fill_gradient()
ggplot(neighborhoods) + 
  geom_sf(aes(fill = avg_canopy)) +
  scale_fill_gradient(low = "#edf8e9", high = "#005a32")
```


Great! You're making progress in improving the graphics. The new layer type `geom_sf()` is a big help for creating maps in `ggplot2`. Altering the colors made the maps much more readable and you probably noticed that they seem to show a different pattern. How about doing this with `tmap`? See the next exercise.


# Task 6: Create a map using tmap
The `tmap` package is an excellent tool for making maps. You'll see that it simplifies the process of binning your data values for nice maps and makes it easy to stitch together different layers.

`tmap` operates similarly to `ggplot2` in the sense that it starts with a function, in this case `tm_shape()`, to set up the map and then is followed by layers and other instructions separated with the `+`.


```{r tmap, eval=FALSE, echo=T}
# Load tmap library
library(tmap)

# Create a simple map of neighborhoods
tm_shape(neighborhoods) + 
    tm_polygons()

# Create a color-coded map of neighborhood tree density
tm_shape(neighborhoods) + 
    tm_polygons(col = "tree_density")

# Style the tree density map
tm_shape(neighborhoods) + 
    tm_polygons("tree_density", palette = "Greens", 
        style = "quantile", n = 7, 
        title = "Trees per sq. KM")

# Create a similar map of average tree canopy
tm_shape(neighborhoods) + 
    tm_polygons("avg_canopy", palette = "Greens", 
        style = "quantile", n = 7, 
        title = "Average tree canopy (%)")
```

You can see the beauty of `tmap`. It makes nice maps in R relatively easily and provides a lot of flexibility to alter your plot as needed

# Task 7: Use tmap to create a final prettier(?) map
In this exercise you will use `tmap` to create a final map with three map views to help you to compare the green you see in an aerial photo with the tree density and average canopy results you computed by neighborhood. The question you're trying to answer is which measure, tree density or average tree canopy, more closely matches what you can see in the aerial photo.

Note: In `map2`, you will make use of the `bbox` argument to force the view extents to match the aerial photo.


## Instructions 1/4


1. Create and view a map of the `manhattan` aerial image with `tm_shape()` and `tm_raster()`.
```{r nice-map-1, eval=FALSE, echo=T}
# Create a map of the manhattan aerial photo
tm_shape(manhattan) + 
  tm_rgb()

```


2. Create and view a map of the `neighboorhoods` borders with `tm_shape()` and `tm_borders()`.

```{r nb-borders, eval=FALSE, echo=T}
# Create a map of the neighborhood polygons
tm_shape(neighborhoods) + 
  tm_borders(col = "black", lwd = 0.5, alpha = 0.5)
```
