---
title: "Map of Blue Mountain ruins"
author: "Sofia Navarro"
date: "2/3/2021"
output: 
  html_document:
    collapsed: FALSE
    code_folding: hide
---

Here is my rendition of an interactive map of the ruins marked by the researcher. The map includes the option of two different views: aerial and physical. This was done using the esri backgrounds. The map also offers a minimap in the bottom right corner. 
Lastly, by clicking on each point, information concerning description, FeatureID and FeatureType of the ruin can be seen.
Click on the "code" button to show the code used to make the map.


```{r setup, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(tidyverse)
```

```{r load, warning=FALSE}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(tidyverse)
```

```{r load, warning=FALSE}
ruins <- read.csv("RCFeature.csv")

esri <- grep("^Esri", providers, value = TRUE)

ruins$Longitude

leaflet() %>%
  addProviderTiles("Esri.WorldPhysical", group = "Physical") %>% 
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  setView(lng=150.2519993, lat =-33.4057078, zoom = 12) %>% 
  addCircleMarkers(clusterOptions = markerClusterOptions(),
                   lng = ruins$Longitude, 
                   lat = ruins$Latitude,
                   popup = paste(ruins$Description, "<br>", ruins$FeatureID, "<br>", ruins$FeatureType)) %>% 
  addLayersControl(
  baseGroups = c("Aerial", "Physical"),
  options = layersControlOptions(collapsed = T)) %>% 
addMiniMap(tiles = "OpenTopoMap", toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
   htmlwidgets::onRender("
                         function(el, x) {
                         var myMap = this;
                         myMap.on('baselayerchange',
                         function (e) {
                         myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                         })
                         }")
    
```
```
